---
title: "Codebook"
author: "Gustavo Niño"
date: "22/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages Used

For the development of the current project where used the following libraries:

```{r}
library(plyr)
library(dplyr)
```

## Data Source

The data was downloaded from the following link: <https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip>
Some additional Information about the data can be found here: <http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones>

```{r}
if(!file.exists("./w4project")){if(!file.exists("./UCI HAR Dataset")){dir.create("./w4project")}}
if(file.exists("./w4project")){setwd("./w4project")}
w4purl<-"https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
if(!file.exists("./w4datasets.zip")){download.file(w4purl, "./w4datasets.zip", method="curl")}
content<-list.files(path=getwd(), pattern= "*.zip", full.names = TRUE)
ldply(.data=content, .fun = unzip, exdir =getwd())
pathuci<-file.path(getwd(), "/UCI HAR Dataset")

```

From the available data, in this project where used the following files:

###   Training Datasets:
    train/X_train.txt
    train/y_train.txt
    train/subject_train.txt

###   Test Datasets:
    test/X_test.txt
    test/y_test.txt
    test/subject_test.txt

###   Other Data Properties
    activity_labels.txt
    features.txt
    
```{r}
files<-list.files(path=pathuci, full.names = TRUE, recursive=TRUE)

train_set<-read.table(files[27], header = FALSE)
train_labs<-read.table(files[28], header = FALSE)
test_set<-read.table(files[15], header = FALSE)
test_labs<-read.table(files[16], header = FALSE)
sub_test<-read.table(files[14], header = FALSE)
sub_train<-read.table(files[26], header = FALSE)
act_labs<-read.table(files[1], header = FALSE)
feats<-read.table(files[2], header = FALSE)
```

(For a deeper understanding of the features.txt content it is adviced to take a look into features_info.txt file). To explore the data loaded there was used the 'str()' function, but wont be showed in this file.

## 0. Label Variables

Before the merging took place, there where labeled the subject and activity variables, to allow a better combination of the data frames in the following steps of the data processing.

```{r}
names(sub_test) = c("subject")
names(sub_train) = c("subject")
x<-as.list(feats$V2)
names(test_set) = x
names(train_set) = x
colnames(act_labs)[2]<-"activity"   
```

(Note that the 'colnames(act_labs)[2]<-"activity"' is part of step 2.)

## 1.  Merges the training and the test sets to create one data set.

For this step first there where created one dataset with training data, and a second dataset with test data, by means of 'cbind()' function, then they where combined by 'rbind()' function.

```{r}
training_ds<-cbind(sub_train, cbind(train_set, train_labs))
test_ds<-cbind(sub_test, cbind(test_set, test_labs))
joined_ds<-rbind(test_ds, training_ds)
```

## 2. Uses descriptive activity names to name the activities in the data set

```{r}
joined_ds<-merge(joined_ds, act_labs, by.x=563, by.y = 1)
joined_ds<-joined_ds[2:564]

```

## 3. Extracts only the measurements on the mean and standard deviation for each measurement. 

```{r}
extracted_ds<- select(joined_ds, "subject", contains("mean")|contains("std"), "activity")
```

## 4. Appropriately labels the data set with descriptive variable names. 

```{r}
names(joined_ds) = gsub("^t", "time_", names(joined_ds))
names(joined_ds) = gsub("^f", "frecuency_", names(joined_ds))
names(joined_ds) = gsub("Acc", "acceleration_", names(joined_ds))
names(joined_ds) = gsub("Body", "body_", names(joined_ds))
names(joined_ds) = gsub("body_body_", "body_", names(joined_ds))
names(joined_ds) = gsub("Gravity", "gravity_", names(joined_ds))
names(joined_ds) = gsub("Gyro", "gyro_", names(joined_ds))
names(joined_ds) = gsub("Jerk", "jerk_", names(joined_ds))
names(joined_ds) = gsub("Mag", "magnitude_", names(joined_ds))
names(joined_ds) = gsub("Gyro", "gyro_", names(joined_ds))
```


## 5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

```{r}
indy_tds<-aggregate(joined_ds[, 2:562],list(c(joined_ds$subject), joined_ds$activity), mean)
colnames(indy_tds)[1]<-"subject"
colnames(indy_tds)[2]<-"activity"
```

## DATA SET FILE CREATION:

```{r}
write.table(joined_ds, file="UCI HAR Dataset/tidy_data.txt", row.names=FALSE)
```



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
